---
title: "Markowit'z Portfolio Theory Mean-Variance Optimization"
author: "Eugene Jang"
date: "2025-06-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Textbook-related library
# library(FRAPO)

# Install Rtools 4.4 
# https://cran.rstudio.com/bin/windows/Rtools/rtools44/rtools.html
# install.packages("quadprog", dependencies = TRUE)
# install.packages(c("ROI.plugin.quadprog", "ROI.plugin.glpk"), dependencies = TRUE)
```

```{r}
library(tidyverse)
library(tidyquant)
library(zoo)
library(xts)
library(PerformanceAnalytics)
library(PortfolioAnalytics)
library(quantmod)
library(timetk)
library(tibbletime)
library(quadprog)
library(ROI)
library(TTR)
```

Example stock and price history

```{r}
apple_df <- getSymbols('AAPL', src='yahoo', auto.assign=FALSE)

# Preview Price/Volume history (6 month period)
chartSeries(apple_df, name="AAPL", subset="last 6 months", theme=chartTheme("white"))
```
Optimal Weights for example portfolio (Minimum Variance)

```{r}
# Create a base vector of ticker symbols
symbols <- c("SPY", "AAPL", "JNJ", "JPM", "XOM", "MSFT")

# Load data from 2015 to today
# Specify the "to = " argument to specify an end date
prices <- getSymbols(
  Symbols = symbols,
  src = "yahoo",
  from = "2015-12-31",
  auto.assign = TRUE,
  warnings = FALSE
) %>%
  # The map function takes an anonymous function and will return a list of five
  # The function Ad() extracts the daily adjusted price series
  map(.f = ~ Ad(get(x = .x))) %>%
  # Use reduce() to merge the elements of .x interactively
  reduce(.f = merge) %>%
  # Use a replacement function to set column names to ticker symbols
  # This function is in prefix form
  # It is equivalent to colnames(x = prices) <- value
  `colnames<-`(value = symbols)
```

```{r}
# Keep only the last reading of each month
# We could have chosen to keep only the first reading of each month
asset_returns_xts <- to.monthly(
  x = prices,
  drop.time = TRUE,
  indexAt = "lastof",
  OHLC = FALSE
) |>
  # Compute simple returns
  # Log returns are time-additive but not portfolio additive
  Return.calculate(method = "discrete") |>
  # Drop the first row since we lose 12/31/2015
  na.omit()
# Keep only the xts returns, ticker symbols, and the prices series
rm(list = setdiff(x = ls(), y = c("symbols", "prices", "asset_returns_xts")))
```

```{r}
# Examine the monthly simple returns for our five ETF's
head(x = asset_returns_xts, 5)
```



```{r}
# Create Portfolio object which is essentially a list object
min_var_portfolio <- portfolio.spec(assets = symbols)
typeof(min_var_portfolio)
```

```{r}
# Add the full investment constraint that specifies that the weights must sum to 1
min_var_portfolio <- add.constraint(
  portfolio = min_var_portfolio,
  type = "full_investment"
)
# Examine the constraint element by extracting min_var_portfolio[["constraints"]][[1]]
str(pluck(.x = min_var_portfolio, "constraints", 1))
```

```{r}
# Add the box constraint that ensure the weights are between 0.1 and 0.6
min_var_portfolio <- add.constraint(
  portfolio = min_var_portfolio,
  type = "box", min = 0.05, max = 0.6
)
# Examine the constraint element by extracting min_var_portfolio[["constraints"]][[2]]
str(pluck(.x = min_var_portfolio, "constraints", 2))
```

```{r}
# Add objective to minimize variance
min_var_portfolio <- add.objective(
  portfolio = min_var_portfolio,
  # Minimize risk
  type = "risk",
  # A character corresponding to a function name, var()
  name = "var"
)
```

```{r}
# Optimization
global_min_portfolio <- optimize.portfolio(
  R = asset_returns_xts,
  portfolio = min_var_portfolio,
  # This defaults to the "quadprog" solver
  optimize_method = "quadprog",
  # Return additional information on the path or portfolios searched
  trace = TRUE
)

# Examine returned portfolio list object
global_min_portfolio
```

Optimal Weights for example portfolio (Maximum Expected Return)

```{r}
# Create Portfolio object
max_exp_return_portfolio <- portfolio.spec(assets = symbols)
```

```{r}
# Add the full investment constraint that specifies the weights must sum to 1
max_exp_return_portfolio <- add.constraint(
  portfolio = max_exp_return_portfolio,
  type = "full_investment"
)
# Add the box constraint that ensure the weights are between 0.1 and 0.6
max_exp_return_portfolio <- add.constraint(
  portfolio = max_exp_return_portfolio,
  type = "box", min = 0.05, max = 0.6
)
```

```{r}
# Add objective to maximize mean returns
max_exp_return_portfolio <- add.objective(
  portfolio = max_exp_return_portfolio,
  # Maximize expected returns
  type = "return",
  # A character corresponding to a function name, mean()
  name = "mean"
)
```

```{r}
# Run the optimization
global_max_portfolio <- optimize.portfolio(
  R = asset_returns_xts,
  portfolio = max_exp_return_portfolio,
  # This defaults to the "glpk" solver
  optimize_method = "ROI",
  # Return additional information on the path or portfolios searched
  trace = TRUE
)
# Examine returned portfolio list object
global_max_portfolio
```
Monthly Portfolio Returns by xts 

```{r}
# Troubleshoot
weights <- extractWeights(global_max_portfolio) # Named numeric vector match names of column names for global portfolio object
names(weights) <- colnames(asset_returns_xts)
length(weights) == ncol(asset_returns_xts)
```

```{r}
# Compute monthly portfolio returns
portfolio_returns_xts_rebalanced_monthly <-
  Return.portfolio(
    R = asset_returns_xts,
    weights = weights,
    # Monthly re-balancing
    rebalance_on = "months",
    # Use simple/arithmetic chaining to aggregate returns
    geometric = FALSE
  ) |>
  `colnames<-`("Monthly_portfolio_returns")

# Examine 
head(portfolio_returns_xts_rebalanced_monthly, 5)
```

Build Portfolio

```{r}
# Set optimal weights
weights <- pluck(.x = global_max_portfolio, "weights")
# Check if the weights and symbols align
tibble(weights, symbols)
```

```{r}
# Ensure that the weights vector sums up to 1
tibble(weights, symbols) |>
  summarize(total_weight = sum(weights))
```

Technical Indicators (Daily)

```{r}
# Simple Moving Average 20 (SMA20)
# Relative Strength Index 14 (RSI14)
# Moving Average Convergence and Divergence

# Extract symbols of adjusted prices
colnames(prices) <- symbols

# Rate of Change (1-day returns)
roc1 <- ROC(prices, n = 1, type = "continuous")
lag_roc1 <- lag.xts(roc1, k = 1) # Lagged log returns (t-1)
target_return <- lag.xts(roc1, k = -1) # Target log future return (t+1)

# SMA(5) applied to each column
sma5 <- do.call(merge, lapply(prices, SMA, n = 5))
colnames(sma5) <- paste0(colnames(prices), "_SMA5")

# RSI(14) applied to each column
rsi14 <- do.call(merge, lapply(prices, RSI, n = 14))
colnames(rsi14) <- paste0(colnames(prices), "_RSI14")

# MACD(Fast:12, Slow:26, Sig:9) 
# Trend-following momentum indicator
macd_list <- lapply(prices, function(x) MACD(x, nFast = 12, nSlow = 26, nSig = 9))

macd_merge <- do.call(merge, lapply(macd_list, function(x) x[,1]))
colnames(macd_merge) <- paste0(symbols, "_MACD")
```

```{r}

```

Risk Performance Features
  Optional static features but helpful for EDA
  
```{r}
# Sharpe Ratio (annualized) 
# Measures risk-adjusted return
sharpe_ratio <- apply(asset_returns_xts, 2, SharpeRatio.annualized)

# Volatility
# Standard deviation of monthly returns
volatility <- apply(asset_returns_xts, 2, sd)

# Mean return per asset
mean_ret <- colMeans(asset_returns_xts)
```

Combine all technical indicators and lagged ROC into feature matrix
```{r}
# Prepare lagged features for feature matrix
features_daily_xts <- merge(
  lag_roc1, 
  sma5, 
  rsi14,
  macd_merge
)

# Clean NA
features_xts <- na.omit(features_xts)

# Convert xts object to matrix x
x <- coredata(features_xts)

# Align and extract target variable y
y <- coredata(target_return[index(features_xts)])

```

Combine features and target variables into Dataframe
```{r}
# Transform matrix into df
feature_matrix <- as.data.frame(x)

# Ensure datatype for return in target variable y is numeric
feature_matrix$prices <- as.numeric(y)

# Normalize 
# feature_matrix <- scale(feature_matrix)
```

